Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-11 paper.tex
updated abstract

% Put your stuff here
%\twocolumn[
%\begin{@twocolumnfalse}
\maketitle
\begin{abstract}
We look to achieve strong scaling in parallelizing numerical ODE methods. We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs. We measure performance and compare in terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
light speedups in our small scale tests up to 64 processors on Harvard's
Odyssey cluster. We analyze possible reasons as to why our implementation
may not be ideal and some tradeoffs that are taken in the Parareal
implementation. Further optimizations are proposed and rationalized as next
steps including ports of to C++ using BLAS. Our code is written in python using mpi4py. We compare tests from various data sources from basic exponential ODEs to larger grids of ODEs to solve for in applications like diffusion and sound wave propagation.

@@

In general, we were unable to see the promised theoretical speedups and error
convergence for the Parareal algorithm. We believe this could be because of
several factors. First, we may have some subtle bugs in our implementation of
the Parareal algorithms. Additionally, we only tested our implementation on
relatively simple problems.  he problem with this testing approach is that the
fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.
We did not see the running time of the Parareal beat that for the serial
implementation of $g_{\textnormal{fine}}$. If we tested our implementation for
more expensive $g_{\textnormal{fine}}$ and more complicated problems, we might
be able to see to see more speedup and better error convergence.

>>>>

% Put your stuff here
%\twocolumn[
%\begin{@twocolumnfalse}
\maketitle
\begin{abstract}
We look to achieve strong scaling in parallelizing numerical ODE methods. We
test both parallelism by result space as well as parallelism by time, with the
Parareal algorithm, to solve various ODEs. We measure performance and compare in
terms of accuracy, speedup and efficiency. We provide some theoretical analysis
for the Parareal algorithm's runtime, speedup, error convergence, and stability.
For the Parareal agorithm, we observe preliminary promising but not convincing
results in our small scale tests up to 32 processors on Harvard's Odyssey
computing cluster. We analyze possible reasons as to why our implementation may
not be ideal and some tradeoffs that are taken in the Parareal implementation.
Further optimizations are proposed and rationalized as next steps including
ports of to C++ using BLAS. Our code is written in python using mpi4py. We
compare tests from various data sources from basic exponential ODEs to larger
grids of ODEs to solve for in applications like diffusion and sound wave
propagation.

@@

In general, we were unable to see the promised theoretical speedups and error
convergence for the Parareal algorithm. We believe this could be because of
several factors. First, we may have some subtle bugs in our implementation of
the Parareal algorithms. Additionally, we only tested our implementation on
relatively simple problems. The problem with this testing approach is that the
fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.
We did not see the running time of the Parareal beat that for the serial
implementation of $g_{\textnormal{fine}}$. If we tested our implementation for
more expensive $g_{\textnormal{fine}}$ and more complicated problems, we might
be able to see to see more speedup and better error convergence.
